---
title: "dpc_exploratory_easy"
author: "Joe Simeone"
date: "2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##library statements
library("tidyverse")
library("readxl") 
library("writexl") 
library("janitor")
library("plotly")
library("knitr")
library("kableExtra")

##Import

dpc_clean <- read_excel("/Users/joesimeone/Desktop/R Master/DPC/clean data/DPC_Tidy_2022_EST.xlsx")

```



```{r, include=FALSE}

##Here, we create 2 pretty similar functions so we can visualize some basic demographic info.
##Allows us to eyeball data as needed for comparison

##County level demographic stratification
strat_county_yr <- function(YrStrat, by_var, expr, StratVar) {
  
dpc_clean %>% 
filter(year==YrStrat) %>% 
group_by(across({{ by_var }})) %>% 
summarise("sum_{{expr}}" := sum({{ expr }})) %>% 
ggplot(aes(county, sum_population_est, fill= {{StratVar}})) +
geom_col(position = "dodge") + scale_y_continuous(labels = scales::comma)}

##Function Calls
county_sex_graph <- strat_county_yr(2022, c(county, sex), population_est, sex) + ylab("Count")
county_race_graph <- strat_county_yr(2022, c(county, race), population_est, race) + ylab("Count")
county_agegr_graph <- strat_county_yr(2022, c(county, age_group_tenyr), population_est, age_group_tenyr) + ylab("Count")


##Allows to look at population composition of Delaware over time 
strat_state_yr <- function(YrStrats, by_var, expr, StratVar){dpc_clean %>% 
    mutate(yearNUM = as.numeric(year)) %>% 
  filter(yearNUM %in% YrStrats) %>% group_by(across({{ by_var }})) %>% 
    summarise("sum_{{expr}}" := sum({{ expr }})) %>% 
    ggplot(aes(year, sum_population_est , fill={{StratVar}})) +
    geom_col() + scale_y_continuous(labels = scales::comma)}

##Function Calls
sex_over_time <- strat_state_yr(2010:2022, c(year, sex), population_est, sex) + ylab("Count")
county_over_time <- strat_state_yr(2010:2022, c(year, county), population_est, county) + ylab("Count")
race_over_time <- strat_state_yr(2010:2022, c(year, race), population_est, race) + ylab("Count")
age_over_time <- strat_state_yr(2010:2022, c(year, age_group_tenyr), population_est, age_group_tenyr) + ylab("Count")
```
```{r, include=FALSE}
##So, this step is to manipulate data so that we can make a clean summary table with KNITR.

##Derives County Estimates
county_for_knitr <- dpc_clean %>% mutate(yearNUM = as.numeric(year)) %>% 
                                  filter(yearNUM %in% 2018:2022) %>% 
                                  group_by(yearNUM, county) %>% 
                                  summarise(county_pop_est = sum(population_est))  

  ##Derives State Estimates
state_for_knitr <- dpc_clean %>% mutate(yearNUM = as.numeric(year)) %>% 
                                  filter(yearNUM %in% 2018:2022) %>% 
                                  group_by(yearNUM) %>% 
                                  summarise(de_pop_est = sum(population_est)) 



##Joins Data on Year
join_county_st <- state_for_knitr %>% left_join(county_for_knitr, by=c('yearNUM'))

```


```{r, include=FALSE}

knitr::kable(join_county_st, "html", caption = "<center><strong>Table 1.0: Population Estimates, By County, 2018-2022</strong></center>", escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "responsive", position = "left")) %>%
  pack_rows(index = c( "2018" = 3, "2019" = 3, "2020"= 3, "2021"= 3, "2022" = 3)) %>% footnote(general = c("<i>Estimates as of October 2022</i>"), escape = FALSE)

```

Purpose: To eyeball Delaware Population Consortium (DPC) estimates stratified by relevant demographic categories. Gives you a quick sense of how population estimates break down.

### DPC Estimates, Recent History.


### DPC, Total Population By County, 2010-2022
```{r, echo=FALSE}
ggplotly(county_over_time)

```



### DPC, Total Population By Sex, 2010-2022
```{r, echo=FALSE}
ggplotly(sex_over_time)

```



### DPC, Total Population By Race, 2010-2022
```{r, echo=FALSE}
ggplotly(race_over_time)
```



### DPC, Total Population By Age Group, 2010-2022
```{r, echo=FALSE}
ggplotly(age_over_time)
```

# County Level Stratifications

### DPC, County BY Sex, 2022
```{r, echo=FALSE}
ggplotly(county_sex_graph)
```



### DPC, County By Race, 2022 
```{r, echo=FALSE}
ggplotly(county_race_graph)
```


### DPC, County By Age Group, 2022
```{r, echo=FALSE}
ggplotly(county_agegr_graph)
```

